---
title: "Data analysis"
author: PengYangYu
date: "2022-11-14"
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
geometry: margin=2.54cm
fontsize: 12pt
mainfont: Times New Roman
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r prepare workspace, include=FALSE, message=FALSE, warning=FALSE}
# Check working directory
getwd()
setwd("E:/things/Duke University/study/2022 Fall/ENVIRON 872/PengYangYu_ENV872_EDA_FinalProject")

# Load packages
library(tidyverse)
library(lubridate)
library(ggplot2)
library(agricolae)
library(sf)
library(mapview)

# Set ggplot theme
mytheme <- theme_bw(base_size = 12) + theme(legend.position = "right")
theme_set(mytheme)

# Load datasets
EPAair_O3 <- read.csv("./Data/Processed/o3_cleaned.csv", 
                      stringsAsFactors = TRUE, row.names = 1) 
EPAair_O3$Date <- ymd(EPAair_O3$Date)
counties_NC_sf<- st_read('./Data/Raw/Spatial/cb_2018_us_county_20m.shp') %>%
  filter(STATEFP == 37)
```

### Research question 1:

Is there any correlation between ozone concentration and daily AQI in 2022?

The null hypothesis: ozone concentration has no correlation with daily AQI

The alternative hypothesis: ozone concentration has correlation with daily AQI

```{r Q1 Linear regression, warning=FALSE, echo=FALSE}
# Filter data frame for data from 2022 
EPAair_O3_22 <- filter(EPAair_O3, Year == '2022')

# Scatter plot of concentration and AQI
O3_AQI_plot <- ggplot(EPAair_O3_22, aes(x = DailyAQI, y = Concentration))+
  geom_point(size=1.2)+
  geom_smooth(method="lm", width=0.8)+
  labs(title='Ozone concentration vs. daily AQI', 
       x='Daily AQI',y='Ozone concentration')
#print(O3_AQI_plot)

# Linear model
o3vsAQI <- lm(data=EPAair_O3_22, Concentration~DailyAQI)
summary(o3vsAQI)

```

The R-squared of this linear model is 0.9099, which means that 90.99% variability in daily air quality index (AQI) is explained by changes in ozone concentration. The degrees of freedom of the model is 9287 = 9289-2 since the number of observations is 9289 and the number of parameters is 2. According to the p-value of slope, which is smaller than 0.05, we reject the null hypothesis, so the correlation between ozone concentration and daily AQI in 2022 is significant.

### Research question 2:

Do different sites have equal mean of ozone concentrations in 2022?

The null hypothesis: they have equal mean of concentrations

The alternative hypothesis: they do not have equal mean of concentrations

```{r Q2 ANOVA, echo=FALSE}
# Test for normality
nor = 0
for(site in unique(EPAair_O3_22$SiteName)){
  test <- shapiro.test(EPAair_O3_22$Concentration
                       [EPAair_O3_22$SiteName == site])
  if(test$p.value >= 0.05){nor = nor+1}
}
sprintf("There are %d sites in total and %d of them follow a normal distribution.", 
        length(unique(EPAair_O3_22$SiteName)), nor)

qqnorm(EPAair_O3_22$Concentration)
qqline(EPAair_O3_22$Concentration)

# Test for equal variance among sites
bartlett.test(EPAair_O3_22$Concentration, EPAair_O3_22$SiteName)

# ANOVA
sites_anova <- aov(data = EPAair_O3_22, Concentration ~ SiteName)
summary(sites_anova)

# Extract groupings for pairwise relationships
sites_group <- HSD.test(sites_anova, 'SiteName', group = TRUE)
#print(sites_group)
```

From the Shapiro-Wilk test result and normal Q-Q plot, we find that most of the sites are conform to normal population distribution assumption. However, the Bartlett's test result shows that the null hypothesis that the variances in each sites are the same is rejected. Since ANOVA is robust against departures from equal variance, we can still apply one-way ANOVA on our dataset.

The p-value of the ANOVA is smaller than 0.05, so we reject the null hypothesis. Therefore, the mean of ozone concentrations in 2022 significantly differ among sites.

### Research question 3:

Is the mean of ozone concentrations in 2021 and 2022 equivalent?

The null hypothesis: the mean between 2021 and 2022 is equivalent

The alternative hypothesis: the mean between 2021 and 2022 is not equivalent

```{r Q3 T-test, echo=FALSE}
# Filter dataset for data from 2021 and 2022
EPAair_O3_2122 <- filter(EPAair_O3, Year == '2021' | Year == '2022')

# Frequency polygons of ozone in 2021 and 2022
o3_freq_2122 <- ggplot(EPAair_O3_2122, aes(x=Concentration, color=as.factor(Year)))+
  geom_freqpoly()+
  labs(x='Ozone concentration',y='Count')+
  scale_color_discrete(name="Year")
print(o3_freq_2122)

# Test for normality
# shapiro.test() asks sample size be between 3 and 5000. Let's see the qqplot.
#shapiro.test(EPAair_O3_2122$Concentration[EPAair_O3_2122$Year == '2021'])
#shapiro.test(EPAair_O3_2122$Concentration[EPAair_O3_2122$Year == '2022'])
qqnorm(EPAair_O3_2122$Concentration)
qqline(EPAair_O3_2122$Concentration)

# Test for equal variance between 2021 and 2022
bartlett.test(EPAair_O3_2122$Concentration, EPAair_O3_2122$Year)

# T-test
year_ttest <- t.test(data = EPAair_O3_2122, Concentration ~ Year)
print(year_ttest)
```

The normal Q-Q plot shows that the data has small deviations from normal distribution. The Bartlett's test result shows that the variances between 2021 and 2022 are different. Again, t-test is robust to these. The T-test suggests that the mean of ozone concentrations in 2021 and 2022 is not equivalent with p-value smaller than 0.05.

### Research question 4:

Is there any trend of ozone concentrations in space?

```{r Q4 Spatial analysis, message=FALSE, echo=FALSE}
# For each site, find the mean values of ozone concentrations in 2022
EPAair_O3_22_mean <- EPAair_O3_22 %>%
  group_by(SiteName, Latitude, Longitude) %>%
  summarise(mean_O3 = mean(Concentration),
            mean_AQI = mean(DailyAQI))

# Convert to an sf object
EPAair_O3_22mean_sf <- EPAair_O3_22_mean %>%
  st_as_sf(coords = c('Longitude','Latitude'), 
           crs = st_crs(counties_NC_sf)$epsg)

# Map the sites by mapview
mapview(counties_NC_sf, col.regions = "red", alpha.regions=0.2, 
        layer.name = 'North Carolina')+
  mapview(EPAair_O3_22mean_sf, zcol = 'mean_O3', 
          layer.name="Ozone concentration")

# Map the sites by ggplot
ggplot() +
  geom_sf(data=counties_NC_sf, fill = 'red', alpha=0.2) +
  geom_sf(data=EPAair_O3_22mean_sf, aes(color=mean_O3), size=3) +
  labs(title = "Ozone concentration in NC")+
  scale_color_continuous(name="Ozone concentration")
```

It seems there is no obvious trend between the location of sites and the mean ozone concentration collected from the sites.

### Research question 5:

Have ozone concentrations changed from 2016 to 2022 at Rockwell?

The null hypothesis: the ozone concentration is stationary over time

The alternative hypothesis: the ozone concentration change over time

```{r Q5 Time series analysis, echo=FALSE}
# Filter data frame for site Rockwell
EPAair_O3_rockwell <- filter(EPAair_O3, SiteName == 'Rockwell')

# Line plot showing ozone concentration over time
ggplot(EPAair_O3_rockwell, aes(x = Date, y = Concentration)) +
  geom_line() +
  labs(title = "Ozone concentrations at Rockwell from 2016 to 2022", 
       y = "Ozone concentration") +
  geom_smooth(method="lm")+
  scale_x_date(date_breaks = "1 years", date_labels = "%Y")

# Interpolate dates
Days <- as.data.frame(seq(as.Date("2016-01-01"), as.Date("2022-11-08"), "day"))
colnames(Days) <- "Date"
EPAair_O3_rockwell <- left_join(Days, EPAair_O3_rockwell, by = "Date") %>%
  mutate('Month'= month(Date), 'Year' = year(Date),
         Concentration = zoo::na.approx(Concentration))

# Generate time series object and decompose it
EPAair_O3_rockwell_ts <- ts(EPAair_O3_rockwell$Concentration, 
                            start = c(2016,1,1), frequency = 365)
Rockwell_decomposed <- stl(EPAair_O3_rockwell_ts, s.window = "periodic")
plot(Rockwell_decomposed)

# Seasonal Mann-Kendall test
Rockwell_trend <- Kendall::SeasonalMannKendall(EPAair_O3_rockwell_ts)
summary(Rockwell_trend)
```

The Seasonal Mann-Kendall test is chosen to test monotonic trend because the decomposed figure shows that the time series object has a strong seasonal component. From the result, we accept the null hypothesis since the p-value is greater than 0.05. Therefore, the ozone concentration at Rockwell is stationary from 2016 to 2022.
